---
title: "Assignment 4"
author: "Nickie Mitch"
date: "10/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load Libraries 
```{r load libraries, message=FALSE}
library(osmdata)
library(opentripplanner)
library(tidyverse)
library(sf)
library(ggthemes)
library(ggspatial)
library(sp)
library(stringr)
library(rgeos)
library(tidygeocoder)
```

#### Load Locations 
Sourced from Open Data DC (https://opendata.dc.gov/datasets/hospitals), this data set shows the location of all 17 hospitals in the District of Columbia.
```{r load locations, message=FALSE}
 DC_Hospitals<- st_read(
  "https://opendata.arcgis.com/datasets/6c18bb76d8644bc1bf53cac2d2199564_4.kml")
```

#### Load Street Data
```{r load street data}
DC_plane <- "+proj=merc +lon_0=-150 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

opq(bbox = 'Washington DC USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_xml(file = 'OTP/graphs/default/dc_streets.osm')

dc_street_features <- opq(bbox = 'Washington DC USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_sf()

dc_streets <- dc_street_features$osm_lines %>%
  st_transform(crs = DC_plane)

ggplot(dc_streets) +
  geom_sf() +
  theme_map()
```

### Set Up Open Trip Planner
```{r launch open trip planner, results = 'hide'}
path_otp <- otp_dl_jar("OTP")
path_data <- file.path(getwd(), "OTP")
path_otp <- paste(path_data, "otp.jar",sep = "/")
otp_build_graph(otp = path_otp, dir = path_data, memory = 1024)
otp_setup(otp = path_otp, dir = path_data, memory =1024)
otpcon <- otp_connect()
```

### Create isochrones 
I'm interested in the area within a 10 minute walk and 10 minute drive of each of the 17 hospitals in Washington, DC. 

```{r create isochrones}
iso_10min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = DC_Hospitals, 
                mode = "WALK", cutoffSec = 600) %>%
  st_transform(crs = DC_plane) %>%
  mutate(mode = "walk")

iso_10min_drive <- 
  otp_isochrone(otpcon = otpcon, fromPlace = DC_Hospitals, 
                mode = "CAR", cutoffSec = 600) %>%
  st_transform(crs = DC_plane) %>%
  mutate(mode = "drive")

iso_all_modes <- rbind(iso_10min_drive, iso_10min_walk)

```

#### Figure 1
```{r create map of isochrones}
right_side <- st_bbox(iso_all_modes)$xmax
left_side  <- st_bbox(iso_all_modes)$xmin
top_side <- st_bbox(iso_all_modes)$ymax
bottom_side <- st_bbox(iso_all_modes)$ymin

ggplot(iso_all_modes) +
  annotation_map_tile(zoomin = 0, type = "cartolight", progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.5) +
  geom_sf(data = DC_Hospitals) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(name = "Area that is reachable\nwithin 10 minutes",
                       labels = c("By car", "By foot")) +
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")
```  

### Calcuate and compare isochrone areas

#### Figure 2 
```{r calc and compare areas 1}
iso_areas <- iso_all_modes %>%
  mutate(area = st_area(iso_all_modes)) %>%
  st_set_geometry(NULL) %>%
  pivot_wider(names_from = mode, values_from = area) 

ggplot(iso_areas, 
       aes(x = as.numeric(walk), y = as.numeric(drive))) +
  geom_point() +
  scale_x_continuous(name = 
            "Area within a ten-minute walking distance\nof a hospital\n(square km)",
            breaks = breaks <- seq(0, 2000000, by = 100000),
            labels = breaks / 1000000) +
  scale_y_continuous(name = 
            "Area within a ten-minute driving distance\nof a hospital\n(square km)",
            breaks = breaks <- seq(0, 70000000, by = 2000000),
            labels = breaks / 1000000) +
  theme_bw()
```

### Calcuating Hospital Access 
While access to health care is a complex issue made up of many more components than simply distance to hospitals, travel time to a hospital is a key determinant of outcomes when it comes to health emergencies like strokes and cardiac arrest. To further understand disparities in health access in Washington, DC - given the substantial areas identified in my isochrone map as being more than a 10 minute walk or drive to any hospital -  I calculated a rudimentary "health score" for points across the district based on the number of hospitals within a 10 minute drive.

Points derived from DC Health Planning Neighborhoods statistical boundaries found on Open Data DC. 

#### Figure 3

```{r import and set up nhood}
nhoods <- st_read("https://opendata.arcgis.com/datasets/de63a68eb7674548ae0ac01867123f7e_13.kml", 
                  quiet = TRUE) 
nhoods1 <- as(nhoods, "Spatial")

points <-spsample(nhoods1, n=1000, type='regular')
```

```{r}
plot(points)
```

```{r make dataframe}
points2 <- st_as_sf(x = points,                         
           coords = coords,
           crs = 4326)
```

```{r iso}
iso_10min_drive3 <- 
  otp_isochrone(otpcon = otpcon, fromPlace = points2, 
                mode = "CAR", cutoffSec = 600)
```

```{r otp stop}
otp_stop()
```

```{r transform 2}

```

```{r count hosp}
iso_10min_drive3 <- iso_10min_drive3%>%
  mutate(health_score = lengths(st_covers(geometry, DC_Hospitals)))
```
```{r out}

out <- data.frame(str_split_fixed(iso_10min_drive3$fromPlace, ",", 2))

out <- st_as_sf(x = out,                         
           coords = c("X2", "X1"),
           crs = 4326)
  
out$health_score <- iso_10min_drive3$health_score
```

```{r plot 3}
ggplot(nhoods) +
  geom_sf(fill="NA", color="grey")+
  geom_sf(data = out, aes(color=health_score))+
  scale_color_gradientn(name="Number of hospitals\nwithin a 10 min drive", colors=c("darkred", "yellow", "blue"))+
  theme_map()+
  theme(legend.position = c(.7,0),
        plot.title = element_text(hjust = 0.5))+
  labs(title = "Hospital Acessibility Map")

```


