---
title: "Assignment 1"
author: "Nickie Mitch"
date: "9/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries

```{r load libraries, message=FALSE}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(ggthemes)
```

## Reviewing variables

```{r view variables, message=FALSE}
area_vars_2010 <- load_variables(2010, "sf1")
```

## Selecting and loading variables

``` {r load variables, message=FALSE}
vars <- c(tot_housing_units = "H001001",
          tot_vacant = "H003003",
          tot_one_person_hh = "P019002",
          tot_rented = "H004004",
          tot_urban = "H002002",
          tot_pop = "P003001", 
          tot_female = "P012026",
          tot_male = "P012002")
Barth_tracts <- get_decennial(geography = "tract", variables = vars,
                           state = "IN", county = "Bartholomew",
                           output = "wide") %>%
  mutate(urban = (tot_urban == tot_housing_units) ,
         pct_vacant = tot_vacant / tot_housing_units,
         pct_rented = tot_rented / tot_housing_units,
         majority_gender = case_when(
           tot_female / tot_pop > 0.5 ~ "female",
           tot_male / tot_pop > 0.5 ~ "male",
           TRUE ~ "None")) %>%
  filter(tot_pop > 0, tot_housing_units > 0) %>%
  select(GEOID, urban, majority_gender, pct_rented, pct_vacant, tot_one_person_hh, tot_pop) 

```

## Displaying selected variables

```{r show loaded microdata variables, message=FALSE}
Barth_tracts
```

## Figure 1 - Scatterplot with all variables

```{r generate figure 1}
ggplot(Barth_tracts, 
       aes(x = pct_rented,
           y = tot_pop,
           size = tot_one_person_hh,
           shape = urban,
           color = majority_gender,
           alpha = pct_vacant )) +
  geom_point() +
  scale_size_continuous(name = "Total one-person households in tract") +
  scale_x_continuous(name = "Percent of households in rented homes",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_y_continuous(name = "Total population of tract",
                     breaks = seq(0, 10000, by = 1000),
                     labels = paste(seq(0, 10000, by = 1000), sep = "")) +
  scale_color_discrete(name = "Majority gender",
                       labels = c("Female",
                       "Male")) +
  scale_alpha_continuous(name = "Percent of housing units vacant") +
   scale_shape(name = element_blank(), 
              labels = c("Rural tract", "Urban tract")) +
  theme_bw() +
  theme(legend.key.size = unit(0.1, "cm"))
```

## Figure 2 - Bar chart with all categorical variables

```{r generate figure 2}
ggplot(Barth_tracts, aes(x = urban, fill = majority_gender)) +
  geom_bar() +
  scale_y_continuous(name = "Number of tracts",
                     breaks = seq(0, 10, by = 1),) +
  scale_x_discrete(name = "",
                        labels = c("Rural Tracts",
                                   "Urban Tracts")) +
  scale_fill_discrete(name = "Majority Gender",
                        labels = c("Female",
                                   "Male")) +
  theme_economist() 
```

## Figure 3 - Scatterplot with two continuous variables and one categorical variable

```{r generate figure 3}
ggplot(Barth_tracts, 
       aes(x = pct_vacant,
           y = tot_pop,
           color = urban)) +
  geom_point(size=4) +
  scale_x_continuous(name = "Percent of housing units vacant",
                     breaks = seq(0, 1, by = 0.01),
                     labels = paste(seq(0, 100, by = 1), "%", sep = "")) +
  scale_y_continuous(name = "Total population of tract",
                     breaks = seq(1000, 10000, by = 1000)) +
  scale_color_discrete(name = "Setting",
                       labels = c("urban", "rural")) +
  theme_wsj()
```

## Figure 4 - Scatterplot with two continuous variables and one categorical variable with a trendline


```{r generate figure 4}
ggplot(Barth_tracts, 
       aes(x = pct_vacant,
           y = tot_pop,
           color = urban)) +
  geom_point() +
  stat_smooth(method = "lm") +
  scale_x_continuous(name = "Percent of housing units vacant",
                     breaks = seq(0, 1, by = 0.01),
                     labels = paste(seq(0, 100, by = 1), "%", sep = "")) +
  scale_y_continuous(name = "Total population of tract",
                     breaks = seq(1000, 10000, by = 1000)) +
  scale_color_discrete(name = "Setting",
                       labels = c("rural", "urban")) +
  theme_bw()
```

## Figure 5 - Scatterplot in Polar Coordinate System with Two Continuous Variables and one categorical variable

```{r generate figure 5}
ggplot(Barth_tracts, 
       aes(x = pct_vacant,
           y = tot_pop,
           color = majority_gender)) +
  geom_point() +
  scale_x_continuous(name = "Percent of housing units vacant",
                     breaks = seq(0, 1, by = 0.01),
                     labels = paste(seq(0, 100, by = 1), "%", sep = "")) +
  scale_y_continuous(name = "Total population of tract",
                     breaks = seq(1000, 10000, by = 1000)) +
  scale_color_discrete(name = "Majority Gender",
                       labels = c("female", "male")) +
  coord_polar(theta = "x") +
  theme_bw()
```

## Figure 6 - Scatterplot with two continuous variables

```{r generate figure 6}
ggplot(Barth_tracts, 
       aes(x = pct_rented,
           y = tot_pop,)) +
  geom_point(size = 4) +
  scale_x_continuous(name = "Percent of housing units vacant",
                     breaks = seq(0, 1, by = 0.05),
                     labels = paste(seq(0, 100, by = 5), "%", sep = "")) +
  scale_y_continuous(name = "Total population of tract",
                     breaks = seq(1000, 10000, by = 1000)) +
  theme_bw() +
  theme(legend.key.size = unit(0.1, "cm"))
```

## Figure 7 - Box plot with two categorical variables and one continuous variable
```{r generate figure 7}
ggplot(Barth_tracts, 
       aes(x = majority_gender,
           y = tot_one_person_hh,
           color = majority_gender)) +
  geom_boxplot(size=1) +
  scale_x_discrete(name = "Majority gender",
                       labels = c("Female",
                                  "Male")) +
  scale_y_continuous(name = "Total one person households",
                     breaks = seq(0, 1000, by = 100),
                    labels = paste(seq(0, 1000, by = 100), sep = "")) +
  theme_bw() +
  theme(legend.position = "none")
```

## Figure 8 - Violin plot with two categorical variables and one continuous variable with scatter plot on top
```{r generate figure 8}
ggplot(Barth_tracts, 
       aes(x = majority_gender,
           y = tot_one_person_hh,
           color = majority_gender)) +
  geom_violin() +
  geom_point(position="jitter",
             alpha =0.5,
             size =2) +
  scale_x_discrete(name = "Majority gender in tract",
                       labels = c("Female",
                                  "Male")) +
  scale_y_continuous(name = "Total one person households in tract",
                     breaks = seq(0, 1000, by = 100),
                    labels = paste(seq(0, 1000, by = 100), sep = "")) +
  theme_bw() +
  theme(legend.position = "none")
```

## Figure 9 - Pie chart with one categorical variable
```{r, generate figure 9}
pie_chart <- ggplot(Barth_tracts, aes(x = 1, fill = urban )) +
  geom_bar(position = "fill") +
  coord_polar(theta = "y") +
  scale_y_continuous(name = "Number of tracts") +
  scale_fill_discrete(name = "",
                       labels = c("Rural Tracts", 
                                  "Urban Tracts")) +
  theme_void()  

pie_chart
```

## Figure 10 - Scatterplot with two continuous variables and one categorical variable with a trendline
```{r generate figure 10}
ggplot(Barth_tracts, 
       aes(x = pct_rented,
           y = tot_one_person_hh,
           color = majority_gender)) +
  geom_point() +
  stat_smooth(method = "lm") +
  scale_x_continuous(name = "Percent of housing units rented in tract",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_y_continuous(name = "Total one person households in tract",
                     breaks = seq(0, 1000, by = 100)) +
  scale_color_discrete(name = "Majority Gender",
                       labels = c("Female", "Male")) +
  theme_economist()
```