---
title: "Assignment 2"
author: "Nickie Mitch"
date: "9/16/2020"
output: html_document
---

## Load libraries

```{r load libaries, message=FALSE}
library(tidyverse)
library(ggplot2)
library (ggthemes)
```

## Load Household-level Data
```{r load hh data, message=FALSE}
hh_data <- read_csv("household_data_me.csv")
attach(hh_data)
```

## Distribution of household income over the past 12 months (HINCP)

This is a continuous variable.

### Mean

```{r HINCP mean, message=FALSE}
paste("The sample mean for household income over the last 12 months is",
      prettyNum(mean(HINCP)), "dollars in USD, using ADJINC to adjust to constant dollars.")
```
### Median

```{r HINCP median, message=FALSE}
paste("The sample median for household income over the last 12 months is",
      prettyNum(median(HINCP)), "dollars in USD, using ADJINC to adjust to constant dollars.")
```

### Standard Deviation
```{r HINCP standard deviation, message=FALSE}
paste("The sample standard deviation for household income over the last 12 months is",
      prettyNum(sd(HINCP)), "dollars in USD, using ADJINC to adjust to constant dollars.")
```

### Interquartile Range 
```{r HINCP interquartile range, message=FALSE}
paste("The sample interquartile range for household income over the last 12 months is",
      quantile(HINCP)["25%"], "to", quantile(HINCP)["75%"], "dollars in USD, using ADJINC to adjust to constant dollars.")
```

### 95% Confidence Interval
```{r HINCP T-Test, message=FALSE}
pop_mean_HINCP <- t.test(HINCP)
pop_mean_HINCP
```

```{r HINCP 95 confidence interval, message=FALSE}
paste("The 95-percent confidence interval for the population mean for household income over the last 12 months is",
       prettyNum(pop_mean_HINCP$conf.int[1]), "to", prettyNum(pop_mean_HINCP$conf.int[2]), "dollars in USD, using ADJINC to adjust to constant dollars.")
```
### Histogram
```{r fancy histogram HINCP, message=FALSE}
ggplot(hh_data, aes(x = HINCP)) +
  geom_histogram() +
  theme_economist() +
  scale_x_continuous(name = "Household income over the last 12 months") +
  scale_y_continuous(name = "Number of households in sample") + 
  annotate("text", x = 700000, y = 5000, size = 3, 
           label = paste("Population mean (95% confidence):",
                         prettyNum(pop_mean_HINCP$conf.int[1]),
                         "to",
                         prettyNum(pop_mean_HINCP$conf.int[2]),
                         "dollars.",
                         "\nSample mean:", 
                         prettyNum(mean(HINCP)), "dollars",
                         "\nSample median:",
                         prettyNum(median(HINCP)), "dollars",
                         "\nSample standard deviation:",
                         prettyNum(sd(HINCP)), "dollars", 
                         "\nSample interquartile range:",
                         prettyNum(quantile(HINCP)["25%"]),
                         "to",
                         prettyNum(quantile(HINCP)["75%"]),
                         "dollars"),
           hjust = 1, vjust = 1) 
```

## Distribution of number of related children in this household (NRC)

This is a continuous variable.

### Mean

```{r NRC mean, message=FALSE}
paste("The sample mean for number of related children in this household is",
      prettyNum(mean(NRC)), "children.")
```
### Median

```{r NRC median, message=FALSE}
paste("The sample median for number of related children in this household is",
      prettyNum(median(NRC)), "children.")
```

### Standard Deviation
```{r NRC standard deviation, message=FALSE}
paste("The sample standard deviation for number of related children in this household is",
      prettyNum(sd(NRC)), "children.")
```

### Interquartile Range 
```{r NRC interquartile range, message=FALSE}
paste("The sample interquartile range for number of related children in this household is",
      quantile(NRC)["25%"], "to", quantile(NRC)["75%"], "children.")
```

### 95% Confidence Interval
```{r NRC T-Test, message=FALSE}
pop_mean_NRC <- t.test(NRC)
pop_mean_NRC
```

```{r NRC 95 confidence interval, message=FALSE}
paste("The 95-percent confidence interval for the population mean for number of related children in this household is",
       prettyNum(pop_mean_NRC$conf.int[1]), "to", prettyNum(pop_mean_NRC$conf.int[2]), "children.")
```

### Histogram
```{r fancy histogram NRC, message=FALSE}
ggplot(hh_data, aes(x = NRC)) +
  geom_histogram() +
  theme_economist() +
  scale_x_continuous(name = "Number of related chidlren in a household") +
  scale_y_continuous(name = "Number of households in sample") + 
  annotate("text", x = 8, y = 5000, size = 3, 
           label = paste("Population mean (95% confidence):",
                         prettyNum(pop_mean_NRC$conf.int[1]),
                         "to",
                         prettyNum(pop_mean_NRC$conf.int[2]),
                         "children",
                         "\nSample mean:", 
                         prettyNum(mean(NRC)), "children",
                         "\nSample median:",
                         prettyNum(median(NRC)), "children",
                         "\nSample standard deviation:",
                         prettyNum(sd(NRC)), "children", 
                         "\nSample interquartile range:",
                         prettyNum(quantile(NRC)["25%"]),
                         "to",
                         prettyNum(quantile(NRC)["75%"]),
                         "children"),
           hjust = 1, vjust = 1) 
```




## Distribution of number of persons in this household (NP)
This is a continuous variable.

### Mean

```{r NP mean, message=FALSE}
paste("The sample mean for number of persons in a household is",
      prettyNum(mean(NP)), "people.")
```
### Median

```{r NP median, message=FALSE}
paste("The sample median for number of persons in a household is",
      prettyNum(median(NP)), "people.")
```

### Standard Deviation
```{r NP standard deviation, message=FALSE}
paste("The sample standard deviation for number of persons in a household is",
      prettyNum(sd(NP)), "people.")
```

### Interquartile Range 
```{r NP interquartile range, message=FALSE}
paste("The sample interquartile range for number of persons in a household is",
      quantile(NP)["25%"], "to", quantile(NP)["75%"], "people.")
```

### 95% Confidence Interval
```{r NP T-Test, message=FALSE}
pop_mean_NP <- t.test(NP)
pop_mean_NP
```

```{r NP 95 confidence interval, message=FALSE}
paste("The 95-percent confidence interval for the population mean for number of persons in a household is",
       prettyNum(pop_mean_NP$conf.int[1]), "to", prettyNum(pop_mean_NP$conf.int[2]), "people.")
```

### Histogram
```{r fancy histogram NP, message=FALSE}
ggplot(hh_data, aes(x = NP)) +
  geom_histogram() +
  theme_economist() +
  scale_x_continuous(name = "Number of persons in a household") +
  scale_y_continuous(name = "Number of households in sample") + 
  annotate("text", x = 10, y = 5000, size = 3, 
           label = paste("Population mean (95% confidence):",
                         prettyNum(pop_mean_NP$conf.int[1]),
                         "to",
                         prettyNum(pop_mean_NP$conf.int[2]),
                         "people",
                         "\nSample mean:", 
                         prettyNum(mean(NP)), "people",
                         "\nSample median:",
                         prettyNum(median(NP)), "people",
                         "\nSample standard deviation:",
                         prettyNum(sd(NP)), "people", 
                         "\nSample interquartile range:",
                         prettyNum(quantile(NP)["25%"]),
                         "to",
                         prettyNum(quantile(NP)["75%"]),
                         "people"),
           hjust = 1, vjust = 1) 
```



## Proportions of internet access (ACCESS) 

This is a categorical variable.

### List all possible values
```{r list values ACCESS, message=FALSE}
access <- unique(hh_data$ACCESS_label)
access
```

### Proportion and confidence intervals for each value
```{r number proportions and conf ACCESS, message=FALSE}
Yes_1 <- t.test(ACCESS_label == access[1])
No <- t.test(ACCESS_label == access[2])
Yes_2 <- t.test(ACCESS_label == access[3])
shares <- tibble(access = c("Yes, by paying a cell phone company or Internet service provider", 
                            "No access to the Internet at this house, apartment, or mobile home", 
                            "Yes, without paying a cell phone company or Internet service provider"),
                 `Share` = c(Yes_1$estimate, 
                             No$estimate, 
                             Yes_2$estimate),
                 Low = c(Yes_1$conf.int[1], 
                             No$conf.int[1], 
                             Yes_2$conf.int[1]),
                 High = c(Yes_1$conf.int[2], 
                             No$conf.int[2], 
                             Yes_2$conf.int[2]))
knitr::kable(shares, caption = "Proportions and 95-percent confidence intervals")
```

### Bar Chart
```{r bar chart ACCESS, message=FALSE}
ggplot(shares, aes(x = access, y = Share, ymin = Low, ymax = High)) +
  geom_bar(stat = "identity") +
  geom_errorbar(width = 0.5) +
  theme_bw() +
  scale_x_discrete(labels = c("No access to the Internet", 
                              "Yes, by paying ...",
                            "Yes, wihtout paying...")) +
  scale_y_continuous(name = "Share of population (error bars represent\n 95-percent confidence interval)",
                     breaks = c(0, 0.2, 0.4, 0.6, 0.8),
                     labels = c("0", "20%", "40%", "60%", "80%"))
```

## Proportions of detailed household language (HHLANP)

This is a categorical variable.

### List all possible values
```{r list values HHLANP, message=FALSE}
language <- unique(hh_data$HHLANP_label)
language
```

### Proportion and confidence intervals for each value
```{r number proportions and conf HHLANP, message=FALSE}
English_only <- t.test(HHLANP_label == language[1])
French <- t.test(HHLANP_label == language[2])
Japanese <- t.test(HHLANP_label == language[3])
Spanish <- t.test(HHLANP_label == language[4])
Korean  <- t.test(HHLANP_label == language[5])
German <- t.test(HHLANP_label == language[6])
Other_Native_North_American_Langauges <- t.test(HHLANP_label == language[7])
Arabic <- t.test(HHLANP_label == language[8])
Thai <- t.test(HHLANP_label == language[9])
Italian <- t.test(HHLANP_label == language[10])
Chinese <- t.test(HHLANP_label == language[11])
Pennsylvania_German <- t.test(HHLANP_label == language[12])
Urdu <- t.test(HHLANP_label == language[13])
Cebuano <- t.test(HHLANP_label == language[14])
Tagalog <- t.test(HHLANP_label == language[15])
Romanian <- t.test(HHLANP_label == language[16])
Other_Bantu_Languages <- t.test(HHLANP_label == language[17])
Bosnian <- t.test(HHLANP_label == language[18])
Hindi <- t.test(HHLANP_label == language[19])
Portuguese <- t.test(HHLANP_label == language[20])
Mandarin <- t.test(HHLANP_label == language[21])
Cantonese <- t.test(HHLANP_label == language[22])
Greek <- t.test(HHLANP_label == language[23])
Polish <- t.test(HHLANP_label == language[24])
Filipino <- t.test(HHLANP_label == language[25])
Bengali <- t.test(HHLANP_label == language[26])
Somali <- t.test(HHLANP_label == language[27])
Serbian <- t.test(HHLANP_label == language[28])
Russian <- t.test(HHLANP_label == language[29])
Vietnamese <- t.test(HHLANP_label == language[30])
Luthuanian <- t.test(HHLANP_label == language[31])
Farsi <- t.test(HHLANP_label == language[32])
Chamorro <- t.test(HHLANP_label == language[33])
Turkish <- t.test(HHLANP_label == language[34])
Tamil <- t.test(HHLANP_label == language[35])
Finnish <- t.test(HHLANP_label == language[36])
Telugu <- t.test(HHLANP_label == language[37])
Swahili <- t.test(HHLANP_label == language[38])
              shares <- tibble(language = c("English only household",
"French",
"Japanese",
"Spanish",
"Korean",
"German",
"Other Native North American languages",
"Arabic",
"Thai",
"Italian",
"Chinese",
"Pennsylvania German",
"Urdu",
"Cebuano",
"Tagalog",
"Romanian",
"Other Bantu Languages",
"Bosnian",
"Hindi",
"Portuguese",
"Mandarin",
"Cantonese",
"Filipino",
"Bengali",
"Somali",
"Serbian",
"Russian",
"Vietnamese", 
"Luthuanian", 
"Farsi", 
"Chamorro", 
"Turkish", 
"Tamil",
"Finnish",
"Telugu",
"Swahili"),
                 `Share` = c(English_only$estimate,
French$estimate,
Japanese$estimate,
Spanish$estimate,
Korean$estimate,
German$estimate,
Other_Native_North_American_Langauges$estimate,
Arabic$estimate,
Thai$estimate,
Italian$estimate,
Chinese$estimate,
Pennsylvania_German$estimate,
Urdu$estimate,
Cebuano$estimate,
Tagalog$estimate,
Romanian$estimate,
Other_Bantu_Languages$estimate,
Bosnian$estimate,
Hindi$estimate,
Portuguese$estimate,
Mandarin$estimate,
Cantonese$estimate,
Filipino$estimate,
Bengali$estimate,
Somali$estimate,
Serbian$estimate,
Russian$estimate,
Vietnamese$estimate, 
Luthuanian$estimate, 
Farsi$estimate,
Chamorro$estimate, 
Turkish$estimate, 
Tamil$estimate,
Finnish$estimate,
Telugu$estimate,
Swahili$estimate),
                 Low = c(English_only$conf.int[1],
French$conf.int[1],
Japanese$conf.int[1],
Spanish$conf.int[1],
Korean$conf.int[1],
German$conf.int[1],
Other_Native_North_American_Langauges$conf.int[1],
Arabic$conf.int[1],
Thai$conf.int[1],
Italian$conf.int[1],
Chinese$conf.int[1],
Pennsylvania_German$conf.int[1],
Urdu$conf.int[1],
Cebuano$conf.int[1],
Tagalog$conf.int[1],
Romanian$conf.int[1],
Other_Bantu_Languages$conf.int[1],
Bosnian$conf.int[1],
Hindi$conf.int[1],
Portuguese$conf.int[1],
Mandarin$conf.int[1],
Cantonese$conf.int[1],
Filipino$conf.int[1],
Bengali$conf.int[1],
Somali$conf.int[1],
Serbian$conf.int[1],
Russian$conf.int[1],
Vietnamese$conf.int[1], 
Luthuanian$conf.int[1], 
Farsi$conf.int[1], 
Chamorro$conf.int[1], 
Turkish$conf.int[1], 
Tamil$conf.int[1],
Finnish$conf.int[1],
Telugu$conf.int[1],
Swahili$conf.int[1]),
                 High = c(English_only$conf.int[2], 
French$conf.int[2],
Japanese$conf.int[2],
Spanish$conf.int[2],
Korean$conf.int[2],
German$conf.int[2],
Other_Native_North_American_Langauges$conf.int[2],
Arabic$conf.int[2],
Thai$conf.int[2],
Italian$conf.int[2],
Chinese$conf.int[2],
Pennsylvania_German$conf.int[2],
Urdu$conf.int[2],
Cebuano$conf.int[2],
Tagalog$conf.int[2],
Romanian$conf.int[2],
Other_Bantu_Languages$conf.int[2],
Bosnian$conf.int[2],
Hindi$conf.int[2],
Portuguese$conf.int[2],
Mandarin$conf.int[2],
Cantonese$conf.int[2],
Filipino$conf.int[2],
Bengali$conf.int[2],
Somali$conf.int[2],
Serbian$conf.int[2],
Russian$conf.int[2],
Vietnamese$conf.int[2], 
Luthuanian$conf.int[2], 
Farsi$conf.int[2], 
Chamorro$conf.int[2], 
Turkish$conf.int[2], 
Tamil$conf.int[2],
Finnish$conf.int[2],
Telugu$conf.int[2],
Swahili$conf.int[2]))
knitr::kable(shares, caption = "Proportions and 95-percent confidence intervals")
```

### Bar Chart
```{r bar chart language, message=FALSE}
ggplot(shares, aes(y = language, x = Share, xmin = Low, xmax = High)) +
  geom_bar(stat = "identity") +
  geom_errorbar(width = 0.5) +
  theme_bw()+
  scale_y_discrete(labels = c("Pennsylvania German",
"Arabic",
"Japanese",
"Vietnamese",
"Korean",
"German",
"Other Native North American languages",
"English only household",
"Thai",
"Italian",
"Chinese",
"French",
"Urdu",
"Cebuano",
"Tagalog",
"Romanian",
"Other Bantu Languages",
"Bosnian",
"Hindi",
"Portuguese",
"Mandarin",
"Cantonese",
"Filipino",
"Bengali",
"Somali",
"Serbian",
"Russian",
"Spanish", 
"Luthuanian", 
"Farsi", 
"Chamorro", 
"Turkish", 
"Tamil",
"Finnish",
"Telugu",
"Swahili")) +
  scale_x_continuous(name = "Share of population (error bars represent\n 95-percent confidence interval)",
                     breaks = c(0, 0.2, 0.4, 0.6, 0.8),
                     labels = c("0", "20%", "40%", "60%", "80%"))
```